---
title: "RLadies: Souporaba R in Python v RStudiu"
subtitle: "Primerjava R paketa DESeq2 in implementacije v Pythonu (PyDESeq2)"
author: "Žan Kuralt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: true
---

# **DESeq2**

Let's first load R packages
```{r load r libraries, warning=FALSE, error=FALSE, message=FALSE}
library(DESeq2)
library(reticulate)
library(EnhancedVolcano)
```

```{r importdata}
rc <- read.table("raw_counts.tsv", header = TRUE, row.names = 1)
meta <- read.table("metadata.tsv", header = TRUE, row.names = 1)
```

```{r deseq, include=FALSE}
dds <- DESeqDataSetFromMatrix(
  countData = rc,
  colData = meta,
  design = ~treatment
)

keep <- rowSums(counts(dds)) > 10
dds <- dds[keep, ]

dds <- DESeq(dds)

results <- results(dds, alpha = 0.05)
```

# **pydeseq2**

## Select virtual environment where python libraries are installed (by default this is "r-reticulate"). One can install packages using the py_install() function.

The chunk below creates virtual environment and installs python libraries in there.
```{r create venv and install libraries}
if (!any(virtualenv_list() %in% "RLadies")) {
  virtualenv_create(envname = "RLadies", packages = c("pandas", "pydeseq2"))
}
```

```{r activate_venv, warning=FALSE}
use_virtualenv("RLadies")
```

## Load libraries

```{python import libs}
import os

import pandas as pd

from pydeseq2.dds import DeseqDataSet
from pydeseq2.default_inference import DefaultInference
from pydeseq2.ds import DeseqStats
```

## Import objects created in R (above). Filter counts data by the same threshold (> 10) as was used in R implementation *DESeq2*.

```{python prepare py data}
counts_df = pd.DataFrame(r.rc.T) # r["rc"] works as well.
meta = pd.DataFrame(r.meta)

genes_to_keep = counts_df.columns[counts_df.sum(axis = 0) > 10]
counts_df_filtrd = counts_df[genes_to_keep]
```

## **pydeseq2** 

```{python pydeseq2}
inference = DefaultInference(n_cpus=6)

dds = DeseqDataSet(
  counts=counts_df_filtrd,
  metadata=meta,
  design_factors="treatment",
  inference=inference,
  )
  
dds.deseq2()
stat_res = DeseqStats(dds, alpha = 0.05, contrast = ["treatment", "JQ1", "DMSO"])
  
# After running the summary() command, results are stored in the .results_df dataframe.
stat_res.summary()
```
  
## Combine and compare results
  
```{r combine results}
res_py <- DESeqResults(py$stat_res$results_df) # using py$ one can access objects created in python
res_py <- as.data.frame(res_py)
res_py$src <- "pydeseq2"

res_r <- as.data.frame(results)
res_r$src <- "DESeq2"

combined_results <- rbind(res_py, res_r)
```
  
## Vulcano plot
  
```{r vulcano}
plt <- EnhancedVolcano(combined_results,
  lab = rownames(combined_results),
  x = "log2FoldChange",
  y = "padj",
  selectLab = c(""),
  pCutoff = 10e-32,
  FCcutoff = 1,
  pointSize = 2.0,
  labSize = 4.0,
  colAlpha = 0.6
)
plt1 <- plt +
  theme(plot.subtitle = element_blank()) +
  facet_wrap(~src)

plt1
```

## Scatterplot

```{r}
plot(res_r$log2FoldChange, res_py$log2FoldChange)
```
